---
title: "Bank Analysis"
author: "Kuiyu Zhu"
date: "10/23/2020"
output: html_document
---

### Task1
1. Whether an account has a credit card or not  
2. Whether an account has a loan or not  
3. The average balance for the account  
Embed the 3 features mentioned in ONE graph.  

```{r, message=FALSE, warning=FALSE}
# set up python env
library(reticulate)
use_python('/Library/Frameworks/Python.framework/Versions/3.7/bin/python3')
```


```{python, message=FALSE, warning=FALSE}
# import libraries
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
sns.set_style("dark")
import warnings
warnings.filterwarnings("ignore")
```

```{python, message=FALSE, warning=FALSE}
# load in data
acct = pd.read_csv('data/accounts_analytical.csv')
# create average_balance column
acct['aver_balance'] = acct['min_balance']/2 + acct['max_balance']/2
```


```{python, message=FALSE, warning=FALSE}
# average balance for accounts which have credit cards
b1 = np.mean(acct['aver_balance'][acct['credit_cards'] >=1])
# average balance for accounts which do not have credit cards
acct['credit_cards'] = acct['credit_cards'].fillna(0)
b2 = np.mean(acct['aver_balance'][acct['credit_cards'] == 0])
acct['loan_amount'] = acct['loan_amount'].fillna(0)
# average balance for accounts which have a loan
b3 = np.mean(acct['aver_balance'][acct['loan_amount'] !=0])
# average balance for accounts which do not have a loan
b4 = np.mean(acct['aver_balance'][acct['loan_amount'] ==0])
```

```{python, message=FALSE, warning=FALSE}
x = ['have cc', 'have no cc', 'have loan', 'have no loan']
y = [b1, b2, b3, b4]

fig, ax = plt.subplots()
width = 0.5 # the width of the bars 
ind = np.arange(len(y))  # the x locations for the groups
ax.barh(ind, y, width, color="green")
ax.set_yticks(ind+width/6)
ax.set_yticklabels(x, minor=False)
plt.title('Average Balance Bar Plot')
for i, v in enumerate(y):
    ax.text(v + 2.5, i + .03, str(round(v,2)), color='black', fontweight='bold')
plt.show()
```


### Task2

What is the distribution of all loans and what are their characteristics?  


```{python, message=FALSE, warning=FALSE}
# load in data
acct = pd.read_csv('data/accounts_analytical.csv')
# create loan dataset
loan = acct[['loan_date', 'loan_amount', 'loan_payment', 'loan_term', 'loan_status', 'loan_default']]
loan = loan.dropna()
```


```{python, message=FALSE, warning=FALSE}
# plot distrubution of loan amount
fig = plt.figure(figsize=(8,6))
g = sns.distplot(loan['loan_amount'], bins=20, kde=False)
plt.title('Distribution of Loan Amount')
plt.xlabel('Loan Amount')
plt.ylabel('Count')
plt.show()
```

```{python, message=FALSE, warning=FALSE}
# plot distribution of loan payment
fig = plt.figure(figsize=(8,6))
g = sns.distplot(loan['loan_payment'], bins=20, kde=False)
plt.title('Distribution of Loan Payment')
plt.xlabel('Loan Payment')
plt.ylabel('Count')
plt.show()
```


```{python, message=FALSE, warning=FALSE}
# plot loan term
fig = plt.figure(figsize=(8,6))
g = sns.countplot(loan['loan_term'])
g.set_title('Count of Loan Term')
for p in g.patches:
    g.annotate(str(format(int(p.get_height()), ',d')), (p.get_x()+0.3, p.get_height()*1.01))
plt.xlabel('Loan Term')
plt.ylabel('Count')
plt.show()
```

```{python, message=FALSE, warning=FALSE}
# plot loan status
fig = plt.figure(figsize=(8,6))
g = sns.countplot(loan['loan_status'])
g.set_title('Count of Loan Status')
for p in g.patches:
    g.annotate(str(format(int(p.get_height()), ',d')), (p.get_x()+0.37, p.get_height()*1.01))
plt.xlabel('Loan Status')
plt.ylabel('Count')
plt.show()
```

```{python, message=FALSE, warning=FALSE}
# plot loan default
fig = plt.figure(figsize=(8,6))
g = sns.countplot(loan['loan_default'])
g.set_title('Count of Loan Default')
for p in g.patches:
    g.annotate(str(format(int(p.get_height()), ',d')), (p.get_x()+0.37, p.get_height()*1.01))
plt.xlabel('Loan Default')
plt.ylabel('Count')
plt.show()
```

```{python, message=FALSE, warning=FALSE}
# plot loan term based on loan status
fig = plt.figure(figsize=(8,6))
g = sns.countplot(x = 'loan_term', hue='loan_status', data=loan)
g.set_title('Count of Loan Term Based on Loan Status')
for p in g.patches:
    g.annotate(str(format(int(p.get_height()), ',d')), (p.get_x()+0.12, p.get_height()*1.01))
plt.xlabel('Loan Term')
plt.ylabel('Count')
plt.show()
```

```{python, message=FALSE, warning=FALSE}
# plot loan term based on whether it is defaulted
fig = plt.figure(figsize=(8,6))
g = sns.countplot(x = 'loan_term', hue='loan_default', data=loan)
g.set_title('Count of Loan Term if it is defaulted')
for p in g.patches:
    g.annotate(str(format(int(p.get_height()), ',d')), (p.get_x()+0.12, p.get_height()*1.01))
plt.xlabel('Loan Term')
plt.ylabel('Count')
plt.show()
```

### Task3

Is there a relationship between a good or bad loan and the time an account is opened?  
Is there a relationship between a good or bad loan and the time the loan is created?  
Is there a specific set of accounts that seem to be at higher or lower risk of defaulting?  


```{python, message=FALSE, warning=FALSE}
# load in data
acct = pd.read_csv('data/accounts_analytical.csv')
# create loan dataset
loan = acct[['acct_creation_date', 'loan_date', 'loan_amount', 'loan_payment', 'loan_term', 'loan_status', 'loan_default']]
loan = loan.dropna()
```


```{python, message=FALSE, warning=FALSE}
# split "acct_creation_date"
loan['acct_year'] = pd.DatetimeIndex(loan['acct_creation_date']).year
loan['acct_month'] = pd.DatetimeIndex(loan['acct_creation_date']).month
loan['acct_day'] = pd.DatetimeIndex(loan['acct_creation_date']).day
# drop "acct_creation_date"
del loan['acct_creation_date']
# split "loan_date"
loan['loan_year'] = pd.DatetimeIndex(loan['loan_date']).year
loan['loan_month'] = pd.DatetimeIndex(loan['loan_date']).month
loan['loan_day'] = pd.DatetimeIndex(loan['loan_date']).day
# drop "loan_date"
del loan['loan_date']
```


```{python, message=FALSE, warning=FALSE}
# Is there a relationship between a good or bad loan and the time an account is opened? 
# define whether a loan is good/bad - status/default
loan1 = loan[['loan_status', 'loan_default', 'acct_year', 'acct_month', 'acct_day', 'loan_year', 'loan_month', 'loan_day']]
# encoding loan_status, loan_default
loan1['loan_status'] = loan1['loan_status'].map({'current': 0, 'expired': 1})
loan1['loan_default'] = loan1['loan_default'].map({False: 0, True: 1})
# heatmap of correlation
mask = np.triu(np.ones_like(loan1.corr(), dtype=bool))
heatmap = sns.heatmap(loan1.corr(), mask = mask, vmin=-1, vmax=1, annot=True, cmap='BrBG')
heatmap.set_title('Correlation Heatmap', fontdict={'fontsize':12}, pad=12)
heatmap
```


It seems that there are relationships between loan_status and loan_year & acct_year.  
It seems that there is no any strong relationship between loan_default and others.  
Lets explore the relationships between loan_status and loan_year & acct_year.  

```{python, message=FALSE, warning=FALSE}
loan1['loan_status'] = loan1['loan_status'].map({0:'current', 1:'expired'})
# loan_status & loan_year
fig = plt.figure(figsize=(8,6))
g = sns.countplot(x = 'loan_status', hue='loan_year', data=loan1)
g.set_title('Count of Loan Status Based on Year when the Loan Created ')
plt.xlabel('Loan Status')
plt.ylabel('Count')
plt.show()
```

Conclusion: Loans Created in 1994 are most likely to be expired, and loans created in 1998 are less likely to be expired.  

```{python, message=FALSE, warning=FALSE}
# loan_status & acct_year
fig = plt.figure(figsize=(8,6))
g = sns.countplot(x = 'loan_status', hue='acct_year', data=loan1)
g.set_title('Count of Loan Status Based on Year when the Account Opened')
plt.xlabel('Loan Status')
plt.ylabel('Count')
plt.show()
```

Conclusion: Loans, associated with the account opened in 1993 are most likely to be expired, and loans, associated with the account opened in 1997 are less likely to be expired.  


```{r, message=FALSE, warning=FALSE}
acct = read.csv('data/accounts_analytical.csv', header = TRUE)
# Omit rows containing specific column (loan_default) of NA
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}
loan2 <- completeFun(acct, 'loan_default')
# export new data
write.csv(loan2, 'data/loan2.csv')
```

```{python, message=FALSE, warning=FALSE}
# ignore some columns (features)
loan2 = pd.read_csv('data/loan2.csv')
loan2 = loan2[['statement_frequency', 'num_customers', 'credit_cards', 'max_withdrawal', 'min_withdrawal',
'max_balance', 'min_balance', 'pmt_order_ct', 'pmt_order_avg_amt', 'gender', 'birth_date', 'loan_default']]
loan2.info()
```


```{python, message=FALSE, warning=FALSE}
# data cleaning
# fill NA with 0
loan2 = loan2.fillna(0)
# encoding
loan2['statement_frequency'] = loan2['statement_frequency'].map({'monthly': 0, 'weekly': 1, 'after every transaction': 2})
loan2['gender'] = loan2['gender'].map({'F': 0, 'M': 1})
# transfer birth_date to age
loan2['birth_year'] = pd.DatetimeIndex(loan2['birth_date']).year
loan2['age'] = loan2['birth_year'].map(lambda x: 2020 - x)
del loan2['birth_year']
del loan2['birth_date']
# bool target to int
loan2['loan_default'] = loan2['loan_default'].map({False: 0, True: 1})
```

```{python, message=FALSE, warning=FALSE}
# heatmap of correlation
fig = plt.figure(figsize=(10,10))
mask = np.triu(np.ones_like(loan2.corr(), dtype=bool))
heatmap = sns.heatmap(loan2.corr(), mask = mask, vmin=-1, vmax=1, annot=True, cmap='BrBG')
heatmap.set_title('Correlation Heatmap', fontdict={'fontsize':12}, pad=12)
heatmap
```

It seems that there are strong relationships between loan_default and min_balance & min_withdrawal.  

```{python, message=FALSE, warning=FALSE}
loan2['loan_default'] = loan2['loan_default'].map({0: False, 1:True})
# plot loan_default with min_balance
fig = plt.figure(figsize=(10,10))
g = sns.catplot(x="loan_default", y="min_balance", data=loan2)
plt.show()
```

The result is quite straight forward. If the minimum balance of an account is less than 0, its loan will definitely be default.     
```{python, message=FALSE, warning=FALSE}
# plot loan_default with min_withdrawal
fig = plt.figure(figsize=(10,10))
g = sns.catplot(x="loan_default", y="min_withdrawal", data=loan2)
plt.show()
```

The result is also very clear. If the minimum withdrawal is approaching to 0, then the loan is likely to be defaulted.  

### Task4

For the account with the highest number of transacations, make a time series line plot for the behavior of the account over time, including all debits and credits, the different methods, and the different categories.  


```{python, message=FALSE, warning=FALSE}
# load in data
tran = pd.read_csv('data/transactions.csv')
tran.info()
```


```{python, message=FALSE, warning=FALSE}
data = tran[tran['account_id'] == 8261]
data = data.drop(['id', 'account_id', 'bank', 'account','amount' ], axis=1)
#data = data.set_index('date')
tran = data
data = data.dropna()
```


```{r, message=FALSE, warning=FALSE}
library(plotly)
tran = py$tran
data = py$data

# plot
fig1 <- plot_ly(tran, x = ~date, y = ~balance, color= ~type, type = 'scatter', mode = 'lines')

fig1 <- fig1 %>%  layout(title = 'Time Series Plot of Balance of Diff Types', 
                       yaxis = list(title = 'Balance'), 
                       xaxis = list(title = 'date'))
                       
fig1
```

```{r, message=FALSE, warning=FALSE}
fig2 <- plot_ly(data, x = ~date, y = ~balance, color= ~method, type = 'scatter', mode = 'lines')

fig2 <- fig2 %>%  layout(title = 'Time Series Plot of Balance of Diff Methods', 
                       yaxis = list(title = 'Balance'), 
                       xaxis = list(title = 'date'))
                       
fig2
```

```{r, message=FALSE, warning=FALSE}
fig3 <- plot_ly(data, x = ~date, y = ~balance, color= ~category, type = 'scatter', mode = 'lines')

fig3 <- fig3 %>%  layout(title = 'Time Series Plot of Balance of Diff Category', 
                       yaxis = list(title = 'Balance'), 
                       xaxis = list(title = 'date'))
                       
fig3
```


### Task5

Explore the validity of the data for the case whether or not an account has a credit card and whether or not they have associated credit card transactions. Is there anything worth noting?  

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
# load data sets
acct = read.csv('data/accounts_analytical.csv', header = TRUE)
tran = read.csv('data/transactions.csv', header = TRUE)
```

```{r, message=FALSE, warning=FALSE}
# list of account ids which have credit card transactions
list_id <- c(tran$account_id[tran$method == 'credit card'])
list_id <- unique(list_id)
length(list_id)
# list of account ids which have at least one credit card
list_ID <- c(acct$account_id[acct$credit_cards >= 1])
list_ID <- unique(list_ID)
length(list_ID)
# common - list of account id which has at least one credit card and also does credit card transaction.
common = intersect(list_id, list_ID)
length(common)
```



```{r, message=FALSE, warning=FALSE}
x <- c('Num_id_cc_trans', 'Num_id_cc', 'Num_cmn_od')
y <- c(807, 726, 139)
data <- data.frame(x,y)
barplot(data$y,
        main = 'Barplot of Account Count',
        xlab = 'class',
        ylab = 'count',
        ylim = c(0,1000), 
        names.arg = c('Num_id_tran', 'Num_id_cc', 'Num_cmn_id'),
        col = c('yellow','blue', 'green'))
text(0.7, 840, "807")
text(1.9, 766, "720")
text(3.1, 168, '139')
```

There are 139 accounts which have credit cards and also have credit cards transactions.  

